---
title: "STAT 4355 Final Project"
author: "Gabrielle Allin, Ramesh Kanakala, Hyun Guk Yoo"
subtitle: This is an R script with the purpose of analyzing Concrete Compressive Strength
output: html_notebook
---

### load the data and look at the first few rows
```{r}
conc <- read.csv(file = "Concrete_Data_Yeh.csv")
conc <- data.frame((conc1[,1:8]), sqrt(conc1[,9]))
names(conc)[names(conc) == "sqrt.conc1...9.."] <- "csMPa"
head(conc)
```

### Histogram of csMPa
```{r}
library(ggplot2)
ggplot(conc, aes(x=csMPa)) + 
 geom_histogram(aes(y=..density..), colour="#4D66E1", fill="lightblue")+
 geom_density(alpha=.2, fill="#E96CEA") + geom_vline(aes(xintercept=mean(csMPa)),
            color="blue", linetype="dashed", size=1)
```

### Scatterplots for csMPa vs each Predictor
```{r}
library(gridExtra)

s1 <- ggplot(conc, aes(x=cement, y=csMPa)) + 
  geom_point(shape=18, color="#4D66E1")+
  geom_smooth(method=lm,  linetype="dashed",
             color="darkred", fill="#E96CEA")

s2 <- ggplot(conc, aes(x=slag, y=csMPa)) + 
  geom_point(shape=18, color="#4D66E1")+
  geom_smooth(method=lm,  linetype="dashed",
             color="darkred", fill="#E96CEA")

s3 <- ggplot(conc, aes(x=flyash, y=csMPa)) + 
  geom_point(shape=18, color="#4D66E1")+
  geom_smooth(method=lm,  linetype="dashed",
             color="darkred", fill="#E96CEA")

s4 <- ggplot(conc, aes(x=water, y=csMPa)) + 
  geom_point(shape=18, color="#4D66E1")+
  geom_smooth(method=lm,  linetype="dashed",
             color="darkred", fill="#E96CEA")

s5 <- ggplot(conc, aes(x=superplasticizer, y=csMPa)) + 
  geom_point(shape=18, color="#4D66E1")+
  geom_smooth(method=lm,  linetype="dashed",
             color="darkred", fill="#E96CEA")

s6 <- ggplot(conc, aes(x=coarseaggregate, y=csMPa)) + 
  geom_point(shape=18, color="#4D66E1")+
  geom_smooth(method=lm,  linetype="dashed",
             color="darkred", fill="#E96CEA")

s7 <- ggplot(conc, aes(x=fineaggregate, y=csMPa)) + 
  geom_point(shape=18, color="#4D66E1")+
  geom_smooth(method=lm,  linetype="dashed",
             color="darkred", fill="#E96CEA")

s8 <- ggplot(conc, aes(x=age, y=csMPa)) + 
  geom_point(shape=18, color="#4D66E1")+
  geom_smooth(method=lm,  linetype="dashed",
             color="darkred", fill="#E96CEA")

grid.arrange(s1, s2, s3, s4, s5, s6, s7, s8, nrow = 3)
```

### Build Full Model
```{r}
#full model
lm1 <- lm(csMPa~., data = conc)

#summary with parameter coefficients and other metrics
summary(lm1)

#variance
print(paste("Variance: ", sigma(lm1)^2))

#99% confidence interval 
print("99% Confidence Interval: ")
confint(lm1,level=0.99)

#prediction on future observations whose csMPa values are identical to data set
predout <- predict(lm1, conc, interval = "predict", predict.level = .99)
print("Prediction on identical data: ")
predout
```

### Confidence/Prediction Interval Curves for Full Model
```{r}
#merge data points and prediction outcome object
merged <- cbind(conc, predout)

#cement
#confidence interval curve
p0 <- ggplot(merged, aes(cement, csMPa)) + geom_point() + stat_smooth(method = lm)
p0
#prediction interval curve 
p00 <- p0 + geom_line(aes(y = lwr), color = "red", linetype = "dashed")+
geom_line(aes(y = upr), color = "red", linetype = "dashed")

#slag
#confidence interval curve
p1 <- ggplot(merged, aes(slag, csMPa)) + geom_point() + stat_smooth(method = lm)
p1
#prediction interval curve 
p11 <- p1 + geom_line(aes(y = lwr), color = "red", linetype = "dashed")+
geom_line(aes(y = upr), color = "red", linetype = "dashed")

#flyash             
#confidence interval curve
p2 <- ggplot(merged, aes(flyash, csMPa)) + geom_point() + stat_smooth(method = lm)
p2
#prediction interval curve 
p22 <- p2 + geom_line(aes(y = lwr), color = "red", linetype = "dashed")+
geom_line(aes(y = upr), color = "red", linetype = "dashed")

#water             
#confidence interval curve
p3 <- ggplot(merged, aes(water, csMPa)) + geom_point() + stat_smooth(method = lm)
p3
#prediction interval curve 
p33 <- p3 + geom_line(aes(y = lwr), color = "red", linetype = "dashed")+
geom_line(aes(y = upr), color = "red", linetype = "dashed")

#superplasticizer   
#confidence interval curve
p4 <- ggplot(merged, aes(superplasticizer, csMPa)) + geom_point() + stat_smooth(method = lm)
p4
#prediction interval curve 
p44 <- p4 + geom_line(aes(y = lwr), color = "red", linetype = "dashed")+
geom_line(aes(y = upr), color = "red", linetype = "dashed")

#coarseaggregate      
#confidence interval curve
p5 <- ggplot(merged, aes(coarseaggregate, csMPa)) + geom_point() + stat_smooth(method = lm)
p5
#prediction interval curve 
p55 <- p5 + geom_line(aes(y = lwr), color = "red", linetype = "dashed")+
geom_line(aes(y = upr), color = "red", linetype = "dashed")

#fineaggregate        
#confidence interval curve
p6 <- ggplot(merged, aes(fineaggregate, csMPa)) + geom_point() + stat_smooth(method = lm)
p6
#prediction interval curve 
p66 <- p6 + geom_line(aes(y = lwr), color = "red", linetype = "dashed")+
geom_line(aes(y = upr), color = "red", linetype = "dashed")

#age                        
#confidence interval curve
p7 <- ggplot(merged, aes(age, csMPa)) + geom_point() + stat_smooth(method = lm)
p7
#prediction interval curve 
p77 <- p7 + geom_line(aes(y = lwr), color = "red", linetype = "dashed")+
geom_line(aes(y = upr), color = "red", linetype = "dashed")

#grid.arrange(p0, p00, p1, p11, p2, p22, p3, p33, p4, p44, p5, p55, p6, p66, p7, p77, nrow = 8)
```

### Residual Analysis for Full Model
```{r}
#View(lm1)
library(MASS)

#standardized residuals
print("standardized residuals:")
stdres(lm1)
barplot(height = stdres(lm1), names.arg = 1:1030, 
        main = "Standardized Residuals", xlab = "Index", 
        ylab = "Standardized Resid", ylim=c(-4,4))
#Add cutoff values. Either 2 or 3 can be chosen.
abline(h=3, col = "Red", lwd=2)
abline(h=-3, col = "Red", lwd=2)

#studentized residuals
print("studentized residuals:")
studres(lm1)
barplot(height = studres(lm1), names.arg = 1:1030, 
        main = "Studentized Residuals", xlab = "Index", 
        ylab = "Studentized Resid", ylim=c(-5,5))
#Add cutoff values. Either 2 or 3 can be chosen.
abline(h=3, col = "Red", lwd=3)
abline(h=-3, col = "Red", lwd=3)

#R-student residuals
print("R-student residuals:")
RStudent <- rstudent(lm1)
RStudent
barplot(height = RStudent, names.arg = 1:1030, 
        main = "R Student Residuals", xlab = "Index", 
        ylab = "R Student Resid", ylim=c(-5,5))
cor.level <- 0.05/(2*25)
cor.qt <- qt(cor.level, 21, lower.tail=F) 
RStudent> cor.qt
abline(h=cor.qt , col = "Red", lwd=3)
abline(h=-cor.qt , col = "Red", lwd=3)
```

### Diagnostics for Leverage and Influence for Full Model
```{r}
#influential analysis
myInf <- influence.measures(lm1)
myInf
summary(myInf)

library(car)
dfbetasPlots(lm1,intercept=T)
influenceIndexPlot(lm1)

#variance inflation factors
vif(lm1)

#normal probability plot of residuals
qqPlot(lm1)

#plot of the residuals versus the fitted values
par(mfrow=c(1,1))
residualPlot(lm1, type="rstudent", quadratic=F, col = "dodgerblue",
             pch=16, cex=1.5, cex.axis=1.5, cex.lab=1.5)
```
### Forward/Backward/Stepwise Selection
```{r}
#library(tidyverse)
library(caret)
library(MASS)

# Stepwise regression model
print("Stepwise:")
step.model <- stepAIC(lm1, direction = "both", 
                      trace = FALSE)
summary(step.model)

# Forward regression model
print("Forward:")
forw.model <- stepAIC(lm1, direction = "forward", 
                      trace = FALSE)
summary(forw.model)

# Backward regression model
print("Backward:")
back.model <- stepAIC(lm1, direction = "backward", 
                      trace = FALSE)
summary(back.model)

anova(lm1, step.model, forw.model, back.model)
```

### Build Reduced Model
```{r}
#reduced model
lm2 <- lm(csMPa~cement+slag+flyash+water+age, data = conc)

#summary with parameter coefficients and other metrics
summary(lm2)

#variance
print(paste("Variance: ", sigma(lm2)^2))

#99% confidence interval 
print("99% Confidence Interval: ")
confint(lm2,level=0.99)

#prediction on future observations whose csMPa values are identical to data set
predout2 <- predict(lm2, conc, interval = "predict", predict.level = .99)
print("Prediction on identical data: ")
predout2
```

### Confidence/Prediction Interval Curves for Reduced Model
```{r}
#merge data points and prediction outcome object
merged2 <- cbind(conc, predout2)

#cement
#confidence interval curve
p0 <- ggplot(merged2, aes(cement, csMPa)) + geom_point() + stat_smooth(method = lm)
p0
#prediction interval curve 
p00 <- p1 + geom_line(aes(y = lwr), color = "red", linetype = "dashed")+
geom_line(aes(y = upr), color = "red", linetype = "dashed")

#slag
#confidence interval curve
p1 <- ggplot(merged2, aes(slag, csMPa)) + geom_point() + stat_smooth(method = lm)
p1
#prediction interval curve 
p11 <- p1 + geom_line(aes(y = lwr), color = "red", linetype = "dashed")+
geom_line(aes(y = upr), color = "red", linetype = "dashed")

#flyash             
#confidence interval curve
p2 <- ggplot(merged2, aes(flyash, csMPa)) + geom_point() + stat_smooth(method = lm)
p2
#prediction interval curve 
p22 <- p2 + geom_line(aes(y = lwr), color = "red", linetype = "dashed")+
geom_line(aes(y = upr), color = "red", linetype = "dashed")

#water             
#confidence interval curve
p3 <- ggplot(merged2, aes(water, csMPa)) + geom_point() + stat_smooth(method = lm)
p3
#prediction interval curve 
p33 <- p3 + geom_line(aes(y = lwr), color = "red", linetype = "dashed")+
geom_line(aes(y = upr), color = "red", linetype = "dashed")

#age                        
#confidence interval curve
p7 <- ggplot(merged2, aes(age, csMPa)) + geom_point() + stat_smooth(method = lm)
p7
#prediction interval curve 
p77 <- p7 + geom_line(aes(y = lwr), color = "red", linetype = "dashed")+
geom_line(aes(y = upr), color = "red", linetype = "dashed")

#grid.arrange(p0, p00, p1, p11, p2, p22, p3, p33, p7, p77, nrow = 7)
```

### Residual Analysis for Reduced Model
```{r}
#View(lm2)
library(MASS)

#standardized residuals
print("standardized residuals:")
stdres(lm2)
barplot(height = stdres(lm2), names.arg = 1:1030, 
        main = "Standardized Residuals", xlab = "Index", 
        ylab = "Standardized Resid", ylim=c(-4,4))
#Add cutoff values. Either 2 or 3 can be chosen.
abline(h=3, col = "Red", lwd=2)
abline(h=-3, col = "Red", lwd=2)

#studentized residuals
print("studentized residuals:")
studres(lm2)
barplot(height = studres(lm2), names.arg = 1:1030, 
        main = "Studentized Residuals", xlab = "Index", 
        ylab = "Studentized Resid", ylim=c(-5,5))
#Add cutoff values. Either 2 or 3 can be chosen.
abline(h=3, col = "Red", lwd=3)
abline(h=-3, col = "Red", lwd=3)

#R-student residuals
print("R-student residuals:")
RStudent <- rstudent(lm2)
RStudent
barplot(height = RStudent, names.arg = 1:1030, 
        main = "R Student Residuals", xlab = "Index", 
        ylab = "R Student Resid", ylim=c(-5,5))
cor.level <- 0.05/(2*25)
cor.qt <- qt(cor.level, 21, lower.tail=F) 
RStudent> cor.qt
abline(h=cor.qt , col = "Red", lwd=3)
abline(h=-cor.qt , col = "Red", lwd=3)
```

### Diagnostics for Leverage and Influence for Reduced Model
```{r}
#influential analysis
myInf2 <- influence.measures(lm2)
myInf2
summary(myInf2)

library(car)
dfbetasPlots(lm2,intercept=T)
influenceIndexPlot(lm2)

#variance inflation factors
vif(lm2)

#normal probability plot of residuals
qqPlot(lm2)

#plot of the residuals versus the fitted values
par(mfrow=c(1,1))
residualPlot(lm2, type="rstudent", quadratic=F, col = "dodgerblue",
             pch=16, cex=1.5, cex.axis=1.5, cex.lab=1.5)
```
### ANOVA
```{r}
anova(lm1, lm2)
```